---
  title: "Exercise9/10"
  format: html
---

1.  I used the help function to find out more about the data set.

    ```{r}
    data("airquality")

    ?airquality

    ```

2.  I used vis_dat to to check out the data. It does need cleaning.

```{r}
options(repos = c(CRAN = "https://cran.rstudio.com"))
install.packages("visdat")
library(visdat)

vis_dat(airquality)

```

3.  I used solar radiation as the predictor for ozone because it is often associated with ozone levels. Increased sunlight can increase ozone levels because of chemical reactions.

```{r}
airquality_clean <- na.omit(airquality)
model <- lm(Ozone ~ Solar.R, data = airquality_clean)

summary(model)

```

4.  Yes, this model is valid. It has a significant p-value. and the

5.  The R-squared value tells us the proportion of variation in a dependent variable that can be explained by the independent variable(s) in the model. The closer the R-squared value is to one, the better the better fit. The R-value in this model is high, telling us that it could be a good fit for the data.

6.  I used broom: : augment to predict ozone

    ```{r}

    install.packages("broom")
    library(broom)

    a <- augment(model, data = airquality_clean)

    head(a)

    ```

    7.  I used ggplot to plot the actual vs predicted ozone.

    ```{r}

    library(ggplot2)

    ggplot(a, aes(x = Ozone, y = .fitted)) +
      geom_point(color = "blue") + 
      geom_abline(intercept = 0, slope = 1, color = "red") +  
      labs(
        title = "Actual vs Predicted Ozone Levels",
        x = "Actual Ozone",
        y = "Predicted Ozone",
        subtitle = paste("Correlation:", round(cor(a$Ozone, a$.fitted), 2)) 
      ) +
      theme_minimal()





    ```
