---
title: "dailyassignment22"
output-dir: docs
type: website
format:
  html:
    self-contained: true
---
First, I downloaded streamflow data for the Poudre River

```{r}
library(dataRetrieval)
library(dplyr)
library(lubridate)
library(tsibble)
library(tidymodels)
library(modeltime)


# Example: Cache la Poudre River at Mouth (USGS site 06752260)
poudre_flow <- readNWISdv(siteNumber = "06752260",    # Download data from USGS for site 06752260
                          parameterCd = "00060",      # Parameter code 00060 = discharge in cfs)
                          startDate = "2013-01-01",   # Set the start date
                          endDate = "2023-12-31") |>  # Set the end date
  renameNWISColumns() |>                              # Rename columns to standard names (e.g., "Flow", "Date")
  mutate(Date = yearmonth(Date)) |>                   # Convert daily Date values into a year-month format (e.g., "2023 Jan")
  group_by(Date) |>                                   # Group the data by the new monthly Date
  summarise(Flow = mean(Flow))                       # Calculate the average daily flow for each month
```


```{r}
# Convert to data frame with Date as class 'Date' for compatibility
poudre_df <- poudre_flow |>
  mutate(Date = as.Date(Date))

# Split into training and future dataset
train_df <- poudre_df |>
  filter(Date < as.Date("2023-01-01"))

future_df <- poudre_df |>
  filter(Date >= as.Date("2023-01-01"))

# create a prophet model for forecasting
model_prophet <- prophet_reg() |>
  set_engine("prophet") |>
  fit(Flow ~ Date, data = train_df)

# create an ARIMA model for forecasting
model_arima <- arima_reg() |>
  set_engine("auto_arima") |>
  fit(Flow ~ Date, data = train_df)

# Combine models into a modeltime table
models_tbl <- modeltime_table(
  model_prophet,
  model_arima
)

# forecast the next 12 months
forecast_tbl <- models_tbl |>
  modeltime_forecast(
    new_data = future_df,
    actual_data = train_df
  )
```

I used the code above to forecast the next 12 months of streamflow data in the Poudre River based on last time assignment.
```{r}
# Use Prophet predictions
predicted_df <- forecast_tbl |>
  filter(.model_desc == "PROPHET") |>
  select(Date = .index, Predicted = .value)

# Join with observed
comparison_df <- predicted_df |>
  left_join(future_df, by = "Date") |>
  rename(Observed = Flow) |>
  drop_na()

r2_model <- lm(Observed ~ Predicted, data = comparison_df)
r2_value <- summary(r2_model)$r.squared
r2_value

```
I used the code above to Compute the R2 value between the model predictions and the observed data using a linear model. The R2 value tells us how well a regression model predicts the outcome of a dependent variable. The reported value is fairly high but not as high as other values that have been reported using this dataset, which tellms me that while it is a decent fit it could be better.

```{r}
ggplot(comparison_df, aes(x = Observed, y = Predicted)) +
  geom_point(color = "green", size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Predicted vs Observed Monthly Streamflow",
    subtitle = paste0("R² = ", round(r2_value, 3)),
    x = "Observed Flow",
    y = "Predicted Flow"
  ) +
  theme_minimal()
```
I used the code above to generate a plot of the Predicted vs Observed values and include a 1:1 line, and a linear model line.

